\documentclass[a4paper]{letter}
\usepackage{url}

\begin{document}

\begin{letter}{}
\vspace*{-5.0cm}

\opening{Dear editor:}

Enclosed you will find a manuscript entitled: ``Fast and accurate
protein false discovery rates on large-scale proteomics data sets with
Percolator 3.0''. The manuscript describes the implementation of a set
of new functions in the popular post-processor Percolator that are
helping the method to cope with large datasets. The large-scaled
dataset we use in the article is the draft human proteome as
assembeled by Kim et al 2014. 

We particularly discuss different ways to derive protein level false
discovery rates for such sets, a topic that we think is of great
interest for the field at this point.


The manuscript is an original contribution and none of the work it
describes has been published elsewhere except for in abstract
form. The manuscript will remain with you and will not be submitted
elsewhere, until you made a decision as to its suitability for
publication in Proteomics.

We suggest the following five experts in the analysis of proteomics
data as potential reviewers:

\begin{itemize}

\item Dr. Dirk Valkenborg, Universiteit Antwerpen,\\
\url{dirk.valkenborg@uantwerpen.be}

\item Professor Lennart Martens,
Gent University, \\
\url{lennart.martens@ugent.be}

\item Professor Olga Vitek,
      Northeastern University, \\
\url{o.vitek@neu.edu}

\item Dr. Manour Askenazi, Biomedical Hosting LLC\\
\url{manor@biomedical-hosting.net}

\item Dr Marshall Bern, Protein Metrics, \\
\url{bern@proteinmetrics.com}

\item Professor Bernard Kuster, Technical University of Munich\\
\url{kuster@tum.de}

\end{itemize}

\vspace*{1.5em}

Sincerely,\\[2em]
Mr. Matthew The,\\
Prof. Michael MacCoss,\\
Prof. William S Noble \& \\
Dr. Lukas K\"all

\end{letter}
\end{document}
