\documentclass[a4paper]{letter}
\usepackage{url}

\begin{document}

\begin{letter}{}
\vspace*{-5.0cm}

\opening{Dear editor:}

Enclosed you will find a manuscript entitled, ``Fast and accurate
protein false discovery rates on large-scale proteomics data sets with
Percolator 3.0'', in answer to your call for a special issue in
honour of the Biemann-medalist 2015, Michael MacCoss. The manuscript
describes the implementation of functionality in the popular
post-processor Percolator to enable software to handle large
datasets. The large-scale data set we use in the article is the draft
human proteome as assembled by Kim et al., 2014. The manuscript
discusses different ways to derive protein-level false discovery rates
for large data sets, a topic that we think is of great interest for
the field at this point.

The manuscript is an original contribution, and none of the work it
describes has been published elsewhere except for in abstract form and
as thesis material. The manuscript will remain with you and will not
be submitted elsewhere, until you made a decision as to its
suitability for publication in {\em Journal of The American Society
  for Mass Spectrometry}.

We suggest the following six experts in the analysis of proteomics
data as potential reviewers:

\begin{itemize}

\item Dr. Dirk Valkenborg, Universiteit Antwerpen,\\
\url{dirk.valkenborg@uantwerpen.be}

\item Professor Lennart Martens,
Gent University, \\
\url{lennart.martens@ugent.be}

\item Professor Olga Vitek,
      Northeastern University, \\
\url{o.vitek@neu.edu}

\item Dr. Manour Askenazi, Biomedical Hosting LLC\\
\url{manor@biomedical-hosting.net}

\item Dr Marshall Bern, Protein Metrics, \\
\url{bern@proteinmetrics.com}

\item Professor Bernard Kuster, Technical University of Munich\\
\url{kuster@tum.de}

\end{itemize}

\vspace*{1.5em}

Sincerely,\\[2em]
Mr. Matthew The,\\
Prof. Michael MacCoss,\\
Prof. William S Noble \& \\
Dr. Lukas K\"all

\end{letter}
\end{document}
